# Statistical analysis
Relying on your conceptual model alone, use rTisane to implement a statistical model that assesses the influence of variables you believe to be important (in the context of additional potentially influential factors) on income.

1. Copy and paste your variables and conceptual model from before. 
```{r}
library(rTisane)
# devtools::load_all()
source("conceptualDisambiguation/app.R")

# Add here
```

2. Fill in the query with your variable names. Run the above code snippet before running this one below.
```{r}
# Copy and paste above 
# Replace <iv> with an independent variable of interest
# Replace <dv> with the dependent variable of interest. In this case, the variable for income. 
script <- query(conceptualModel=cm, iv=<iv>, <dv>)
```

3. Copy and paste your statistical modeling script below:  
> Note: You may need to update the path to "/Users/cse-loaner/Git/rTisane/Participants/income_final.csv"
```{r}
# Add statistical model here

```

4. Run the above snippet!

5. Summarize the results of your analysis below: 
